---
title: "Unbinned Analysis"
author: "Emanuele Coradin"
date: "`r Sys.Date()`"
output: 
  read_document: rmdformats::readthedown
  pdf_document:
    number_sections: true
    toc: true
    toc_depth: 2
  html_document:
    number_sections: true
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

color_vector <- c("#CC0000",   # Dark red
                  "#CC79A7",   # Muted purple
                  "#D55E00",   # Vermilion
                  "#009E73",   # Bluish green
                  "#56B4E9",   # Sky blue
                  '#000046',   # Deep Blue
                  "#DB1E60",   # Pinkish-red
                  "#E69F00")   # Yellow-orange

```

```{r, message=FALSE, echo=FALSE}
library(rjags)
library(ggplot2)
library(dplyr)
library(rstan)
library(ggplot2)

```
# Bayesian analysis of the $\mu$ lifetime and g-2 parameter with $\mu$ SR

## Introduction

Muons are long-lived particles, produced in the decays of pions and
kaons originating from the interactions of primary cosmic rays with the
Earth’s atmosphere.

Moreover, parity violation is also present in the decay which proceeds
as follow: $\mu^+ \rightarrow e^+ + \nu_e + \overline{\nu_\mu}$ and
$\mu^- \rightarrow e^- + \overline{\nu_e} + \nu_\mu$.

Simple experiments can be performed by measuring muons that decay in a
thick absorber. If the absorber is immersed in a constant magnetic
field, the muon spin, before the decay, precesses with its Larmor
frequency $\omega_L = g_{\mu} \frac{eB}{2 m_{\mu} c}$.

The decay proceeds mainly along the direction of the spin of the muon
and therefore, if the muon is (partly) polarized, the detected signal
varies with time with $\omega_L$. Analyzing the data collected without
and with the magnetic field, we set up a Markov Chain Monte Carlo that
allows us to extract the muon lifetime $\tau$, with the magnetic field
turned off, and the muon precession frequency with the magnetic field
turned on.

In this document we are presenting an unbinned analysis. See the file about the binned analysis for more details.


## Functions and default values

```{r functions}

# ------- Analytical calculations -----------

getGammaParam <- function(mean, variance) {
  beta <- mean / variance
  alpha <- mean * beta
  c(alpha = alpha, beta = beta)
}

getBetaParam <- function(mean, variance) {
  nu <- mean * (mean * (1 - mean) / variance - 1)
  alpha <- mean * nu
  beta <- nu * (1 - mean)
  c(alpha = alpha, beta = beta)
}

# -------- Functions for plotting -----------

plotdata <- function(data, title = NULL, xlab = expression(Delta * T ~ "[" * mu * "s" * "]"), histdata = NULL, scale = 'lin', bins = bins_default, plot_histbar = FALSE, xlim = NULL, ylim=NULL) {
  if (is.null(histdata)) histdata <- hist(data, breaks = bins, plot = FALSE)

  df <- data.frame(mids = histdata$mids, counts = histdata$counts)

  if (scale == "lin") {
    df <- df %>% mutate(se = sqrt(counts))  # Errore standard come sqrt dei counts (assunzione di distribuzione di Poisson)
  } else {
    df <- df %>% mutate(se = 1 / (log(10)*sqrt(counts)))  # Errore standard come 1/sqrt(counts)
  }

  p <- ggplot() +
    theme_minimal() +
    ggtitle(title) +
    labs(x = xlab, y = ifelse(scale == "lin", "Counts", expression(log[10](Counts))), color = "Legend") +
    theme(plot.title = element_text(face = "bold", size = 14),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12))

  if (plot_histbar) {
    p <- p + geom_bar(data = df, aes(x = mids, y = counts), fill = 'blue', color = 'blue')
    if (scale == "log") {
      p <- p + scale_y_log10()
    }
  } else {
    if (scale == "lin") {
      p <- p +
        geom_point(data = df, aes(x = mids, y = counts), color = 'blue', size = 1) +
        geom_errorbar(data = df, aes(x = mids, ymin = counts - se, ymax = counts + se), color = 'red', width = 0.2, size = 0.5, alpha = 0.6)
    } else {
      p <- p +
        geom_point(data = df, aes(x = mids, y = log10(counts)), color = 'blue', size = 1) +
        geom_errorbar(data = df, aes(x = mids, ymin = log10(counts) - se, ymax = log10(counts) + se), color = 'red', width = 0.2, size = 0.5, alpha = 0.6)
    }
  }

  if (!is.null(xlim)) {
    p <- p + xlim(xlim)
  }
  if (!is.null(ylim)) {
    p <- p + ylim(ylim)
  }

  p
}

fitting <- function(law, parms, data, xlab = expression(Delta * T ~ "[" * mu * "s" * "]"), scale = 'lin', ...) {
  p <- plotdata(data, xlab = xlab, scale = scale, ...)
  if (scale == 'lin') {
    p <- p + geom_line(aes(x = data, y = law(data, parms)), size = 0.8, show.legend = FALSE)
  } else if (scale == 'log') {
    p <- p + geom_line(aes(x = data, y = log10(law(data, parms))), size = 0.8, show.legend = FALSE)
  }
  p
}

plot_posterior_param <- function(samples, param_name, stats, prior_func=NA, xlim=NULL, xlab=NULL,  legend.loc="topright", ...) {
  dens <- density(samples)
  
  if(is.null(xlab))
    xlab <- param_name

  # Calculate the prior density if prior_func is provided
  if(!is.na(prior_func)){
    if(is.null(xlim)) {
      x <- seq(min(samples), max(samples), length = 5000)
    } else {
      x <- seq(xlim[1], xlim[2], length = 5000)
    }
    y_prior <- prior_func(x)
    max_prior <- max(y_prior)
  } else {
    max_prior <- 0
  }

  # Calculate the maximum y value for posterior and prior, then set ylim
  max_posterior <- max(dens$y)
  max_y <- max(max_posterior, max_prior) * 1.1

  plot(dens, main = paste("Posterior Density of", param_name), xlab = xlab, ylab = "Probability Density", col = "skyblue", lwd = 2, xlim=xlim, ylim=c(0, max_y), ...)

  # Overlay the prior density if prior_func is provided
  if(!is.na(prior_func)){
    lines(x, y_prior, col = "darkorange", lwd = 2, lty = 2)
    
    legend(legend.loc, legend = c(
      paste("Mean =", round(stats$mean, 2)),
      paste("SD =", round(stats$sd, 2)),
      paste("2.5% CI =", round(stats$ci[1], 2)),
      paste("97.5% CI =", round(stats$ci[2], 2)),
      "Prior"
      ), 
      col = c("blue", 'lightblue', "red", "red", "darkorange"), lty = c(2, 2, 2, 2), lwd = 2, bty = "n"
    )
  } else {
    legend(legend.loc, legend = c(
      paste("Mean =", round(stats$mean, 2)),
      paste("SD =", round(stats$sd, 2)),
      paste("2.5% CI =", round(stats$ci[1], 2)),
      paste("97.5% CI =", round(stats$ci[2], 2))
      ), 
      col = c("blue", 'lightblue', "red", "red"), lty = c(2, 2, 2), lwd = 2, bty = "n"
    )
  }
  
  abline(v = stats$mean, col = "blue", lwd = 2, lty = 2)
  abline(v = stats$mean + stats$sd, col = "lightblue", lwd = 2, lty = 2)
  abline(v = stats$mean - stats$sd, col = "lightblue", lwd = 2, lty = 2)
  abline(v = stats$ci, col = "red", lwd = 2, lty = 2)
}


extract_stats <- function(summary_obj, param_name=NULL) {
  if(is.null(param_name)){
    list(
      mean = summary_obj$statistics["Mean"],
      sd = summary_obj$statistics["SD"],
      ci = summary_obj$quantiles[c("2.5%", "97.5%")]
    )
  }
  
  else{
    list(
      mean = summary_obj$statistics[param_name, "Mean"],
      sd = summary_obj$statistics[param_name, "SD"],
      ci = summary_obj$quantiles[param_name, c("2.5%", "97.5%")]
    )
  }

}

```

```{r Default values}
bins_default <- 120
bins_lifetime <- 60

# Named vector containing the moments for the priors
prior_moments <- c( #like a dictionary
  tau_mean = 2.1,    tau_var = 0.01^2,
  alpha_mean = 0.06,  alpha_var = 0.01^2,
  omega_mean = 4.77, omega_var = (4.77 * 0.04)^2 / 12,
  delta_mean = 0,    delta_var = 0.03
)

# Named vector containing the parameters of the priors
PriorParam <- list(
  tau = getGammaParam(prior_moments[['tau_mean']], prior_moments[['tau_var']]),        #[[]] selects just the value, not the name
  omega = getGammaParam(prior_moments[['omega_mean']], prior_moments[['omega_var']]),
  delta = getBetaParam((prior_moments[['delta_mean']] + pi/2) / pi, prior_moments[['delta_var']] * pi^-2), #beta is in interval [0,1], so we dilatate and shift the parameters, then we'll reset the function
  alpha = getBetaParam((prior_moments[['alpha_mean']] + 1/3) / (4/3), prior_moments[['alpha_var']] / (4/3)^2)
)

```

## Data Loading and Filtering

```{r }
# Read files
paths_lifetime   <-dir('g-2data/data/lifetime/2023_24', pattern ='^t', full.names=TRUE)
paths_precession <-dir('g-2data/data/precession/2023_24', full.names=TRUE)

data_lifetime   <-numeric()
data_precession <- numeric()

for (path in paths_lifetime)   {data_lifetime   <- as.integer(c(data_lifetime, (readLines(path))))}
for (path in paths_precession) {data_precession <- as.integer(c(data_precession, (readLines(path))))}

cat('Length data_lifetime=',length(data_lifetime),'\n')
cat('Length data_precession',length(data_precession))

# Calibrate data
p0       <- 7.4
sigma_p0 <- 4.4
p1       <- 14.90
sigma_p1 <- 0.11

calibrate <- function(x) (p0 + p1 * x)*1e-3

data_lifetime   <- calibrate(data_lifetime)
data_precession <- calibrate(data_precession)

# Cleaning the dataset
data_lifetime   <- data_lifetime[data_lifetime<8]
#data_precession <- data_precession[data_precession>1 & data_precession<7]

```

## Lifetime Analysis 

```{r}
dataList <- list(x = data_lifetime)

init_values <- list(list(tau = 2.2))

n_burnin <- 2000 # Length of the burn-in phase
thinning <- 1
Nrep     <- 10000 # Number of values to simulate
n_adapt  <- 1000

# Write JAGS model
model_lifetime_def <- sprintf("
model{
  for (i in 1:length(x)) {
    x[i] ~ dexp(1/tau)
  }

  tau ~ dgamma(%.2f, %.2f)
}", PriorParam$tau["alpha"], PriorParam$tau["beta"])

model_lifetime <- jags.model(file = textConnection(model_lifetime_def), data = dataList, inits = init_values, n.chains = 1)

# Adaptation phase
adapt(model_lifetime, n_adapt)

# Burn-in phase
update(model_lifetime, n.iter = n_burnin)

# Sample from the posterior
posterior_lifetime <- coda.samples(model_lifetime, variable.names = c('tau'), n.iter = Nrep, thin = thinning)
(summary_lifetime <- summary(posterior_lifetime))
#plot(posterior_lifetime)

posterior_matrix <- as.matrix(posterior_lifetime)

# Retrieve the chains
tau_samples <- posterior_matrix[, "tau"]

acf(tau_samples, main = "Autocorrelation of tau")
```

## Liferime Analysis Rstan

```{r}
dataList <- list(N = length(data_lifetime), x = data_lifetime)

# Define initial values
init_values <- list(tau = 2.2, a=0.99)

# Save the Stan model to a file
stan_model_code <- "
data {
  int<lower=0> N;      // Number of observations
  real<lower=0> x[N];  // Observations
}

parameters {
  real<lower=0> tau;   // Parameter of interest
  real<lower=0.9, upper=1> a; 
}

model {
  // Priors
  tau ~ gamma(%f, %f);

  // Likelihood with multiplicative additional term
  for (i in 1:N) {
    target += a* exponential_lpdf(x[i] | 1 / tau) + (1-a) * exponential_lpdf(x[i] | 1 / 100);
  }
}
"
write(sprintf(stan_model_code, PriorParam$tau["alpha"], PriorParam$tau["beta"]), file = "model_lifetime.stan")

# Fit the model
fit <- stan(file = "model_lifetime.stan", data = dataList, init = list(init_values), chains = 1, iter = 100, warmup = 10, thin = 1)

# Print the summary of the fitted model
print(fit)
```

```{r }
tau_stats <- extract_stats(summary_lifetime)

prior_tau <- function(x) {
  dgamma(x, shape = PriorParam$tau[["alpha"]], rate = PriorParam$tau[["beta"]])
}

plot_posterior_param(tau_samples, "tau", tau_stats, prior_tau, xlab=expression(tau ~ "[" * mu * "s" * "]"), legend.loc="topleft")

```

## Precession Analysis

```{r}
tau_lifetime_posterior <- getGammaParam(summary_lifetime$statistics[['Mean']] ,summary_lifetime$statistics[['SD']])

```

```{r}

dataList = list(x = data_precession, zeros = rep(0, length(data_precession)), tau=summary_lifetime$statistics['Mean'])

init_values <- list(
  list(delta_base = 0.4, omega = 4.7, alpha_base = 0.05)
)

n_burnin <- 200  # Length of the burn-in phase
thinning <- 1
Nrep     <- 500 # Number of values to simulate
n_adapt  <- 100

model_precession_def <- sprintf("
model {
  for (i in 1:length(x)) {
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <- -L[i]
    L[i] <- -x[i]/tau + log(1+alpha*cos(omega*x[i]+delta)) - log(tau) - log(FACTOR)
  }
  
  delta_base ~ dbeta(%.2f, %.2f) 
  omega ~ dgamma(%.2f, %.2f)
  alpha_base ~ dbeta(%.2f, %.2f) 
  #tau ~ dgamma(%.2f, %.2f)
  
  delta <- delta_base * 3.14 - 3.14/2
  alpha <- alpha_base * 4/3 - 1/3
  
  g <- 0.41986*omega
  
  FACTOR <- 1 + alpha/(1+tau^2*omega^2) * (cos(delta)-tau*omega*sin(delta))
}
", PriorParam$delta["alpha"], PriorParam$delta["beta"], 
   PriorParam$omega["alpha"], PriorParam$omega["beta"],  
   PriorParam$alpha["alpha"], PriorParam$alpha["beta"], 
   tau_lifetime_posterior["alpha"], tau_lifetime_posterior["beta"])


# Create the model
model_precession <- jags.model(file = textConnection(model_precession_def), data = dataList, n.chains = 1)

# Adaptation phase
adapt(model_precession, n_adapt)

# Burn-in phase
update(model_precession, n.iter = n_burnin)


# Sample from the posterior
posterior_precession <- coda.samples(model_precession, variable.names = c('delta', 'omega', 'alpha', 'tau', 'g','FACTOR'), n.iter = Nrep, thin = thinning)
summary_precession <- summary(posterior_precession)
posterior_matrix <- as.matrix(posterior_precession)

# Retrieve the chains
delta_samples <- posterior_matrix[, "delta"]
omega_samples <- posterior_matrix[, "omega"]
alpha_samples <- posterior_matrix[, "alpha"]
tau_samples <- posterior_matrix[, "tau"]
g_samples <- posterior_matrix[, "g"]

```


```{r}
plot(posterior_precession)

par(mfrow = c(2, 2))

acf(delta_samples, main = "Autocorrelation of delta")
acf(omega_samples, main = "Autocorrelation of omega")
acf(alpha_samples, main = "Autocorrelation of alpha")
acf(g_samples, main = "Autocorrelation of g")
```

```{r, eval=FALSE }
# Extract statistics for each parameter
delta_stats <- extract_stats(summary_precession, "delta")
omega_stats <- extract_stats(summary_precession, "omega")
alpha_stats <- extract_stats(summary_precession, "alpha")
g_stats <- extract_stats(summary_precession, "g")

# Define the prior functions for new parameters
prior_delta <- function(x) {
  dbeta((x + 3.14/2) / 3.14, shape1 = PriorParam$delta["alpha"], shape2 = PriorParam$delta["beta"]) / 3.14
}

prior_omega <- function(x) {
  dgamma(x, shape = PriorParam$omega["alpha"], rate = PriorParam$omega["beta"])
}

prior_alpha <- function(x) {
  dbeta((x + 1/3) * 3 / 4, shape1 = PriorParam$alpha["alpha"], shape2 = PriorParam$alpha["beta"]) * 3 / 4
}

# Plotting posterior parameters with priors
#par(mfrow = c(5, 1))  # Set up the plotting area for 5 plots vertically

plot_posterior_param(delta_samples, "delta", delta_stats, prior_delta, xlab = expression(delta ~ '[rad]'))
plot_posterior_param(omega_samples, "omega", omega_stats, prior_omega, xlab = expression(omega ~ '[MHz]' ))
plot_posterior_param(alpha_samples, "alpha", alpha_stats, prior_alpha, xlab = expression(alpha))
plot_posterior_param(g_samples, "g", g_stats, xlab=expression('g' * mu))
```


## Precession Analysis including noise

```{r}
tau_lifetime_posterior <- getGammaParam(summary_lifetime$statistics[['Mean']] ,summary_lifetime$statistics[['SD']])

```

```{r}

dataList = list(x = data_precession, zeros = rep(0, length(data_precession)), tau=summary_lifetime$statistics['Mean'])

init_values <- list(
  list(delta_base = 0.4, omega = 4.7, alpha_base = 0.05, a=0.99)
)

n_burnin <- 5000  # Length of the burn-in phase
thinning <- 2
Nrep     <- 15000 # Number of values to simulate
n_adapt  <- 3000

model_precession_def <- sprintf("
model {
  for (i in 1:length(x)) {
    zeros[i] ~ dpois(zeros.mean[i])
    zeros.mean[i] <- -L[i]
    L[i] <- log( a * exp(-x[i]/tau) * (1+alpha*cos(omega*x[i]+delta)) / (tau * FACTOR) + (1-a) * exp(-x[i]/100) /100 )
  }
  
  delta_base ~ dbeta(%.2f, %.2f) 
  omega ~ dgamma(%.2f, %.2f)
  alpha_base ~ dbeta(%.2f, %.2f) 
  a ~ dunif(0.9,1)
  #tau ~ dgamma(%.2f, %.2f)
  
  delta <- delta_base * 3.14 - 3.14/2
  alpha <- alpha_base * 4/3 - 1/3
  
  g <- 0.41986*omega
  
  FACTOR <- 1 + alpha/(1+tau^2*omega^2) * (cos(delta)-tau*omega*sin(delta))
}
", PriorParam$delta["alpha"], PriorParam$delta["beta"], 
   PriorParam$omega["alpha"], PriorParam$omega["beta"],  
   PriorParam$alpha["alpha"], PriorParam$alpha["beta"], 
   tau_lifetime_posterior["alpha"], tau_lifetime_posterior["beta"])


# Create the model
model_precession <- jags.model(file = textConnection(model_precession_def), data = dataList, n.chains = 1)

# Adaptation phase
adapt(model_precession, n_adapt)

# Burn-in phase
update(model_precession, n.iter = n_burnin)


# Sample from the posterior
posterior_precession <- coda.samples(model_precession, variable.names = c('delta', 'omega', 'alpha', 'tau', 'g','FACTOR', 'a'), n.iter = Nrep, thin = thinning)
summary_precession <- summary(posterior_precession)
posterior_matrix <- as.matrix(posterior_precession)

# Retrieve the chains
delta_samples <- posterior_matrix[, "delta"]
omega_samples <- posterior_matrix[, "omega"]
alpha_samples <- posterior_matrix[, "alpha"]
tau_samples <- posterior_matrix[, "tau"]
g_samples <- posterior_matrix[, "g"]
a_samples <- posterior_matrix[, "a"]

```

```{r }
#plot(posterior_precession)

#par(mfrow = c(3, 2))

acf(delta_samples, main = "Autocorrelation of delta")
acf(omega_samples, main = "Autocorrelation of omega")
acf(alpha_samples, main = "Autocorrelation of alpha")
acf(g_samples, main = "Autocorrelation of g")
acf(a_samples, main = "Autocorrelation of a")
```

```{r }
# Extract statistics for each parameter
delta_stats <- extract_stats(summary_precession, "delta")
omega_stats <- extract_stats(summary_precession, "omega")
alpha_stats <- extract_stats(summary_precession, "alpha")
g_stats <- extract_stats(summary_precession, "g")
a_stats <- extract_stats(summary_precession, "g")

# Define the prior functions for new parameters
prior_delta <- function(x) {
  dbeta((x + 3.14/2) / 3.14, shape1 = PriorParam$delta["alpha"], shape2 = PriorParam$delta["beta"]) / 3.14
}

prior_omega <- function(x) {
  dgamma(x, shape = PriorParam$omega["alpha"], rate = PriorParam$omega["beta"])
}

prior_alpha <- function(x) {
  dbeta((x + 1/3) * 3 / 4, shape1 = PriorParam$alpha["alpha"], shape2 = PriorParam$alpha["beta"]) * 3 / 4
}

# Plotting posterior parameters with priors
#par(mfrow = c(5, 1))  # Set up the plotting area for 5 plots vertically

plot_posterior_param(delta_samples, "delta", delta_stats, prior_delta, xlab = expression(delta ~ '[rad]'))
plot_posterior_param(omega_samples, "omega", omega_stats, prior_omega, xlab = expression(omega ~ '[MHz]' ))
plot_posterior_param(alpha_samples, "alpha", alpha_stats, prior_alpha, xlab = expression(alpha))
plot_posterior_param(g_samples, "g", g_stats, xlab=expression('g'))
plot_posterior_param(a_samples, "a", a_stats, xlab=expression('a'), legend.loc='topleft')
```

